---
title: "Support Vector Machine Assignment"
author: "Michael J. Jones"
date: "23/04/2015"
output: pdf_document
---

# Problem 1
*Suppose that the following are a set of points in two classes:*

`Class 1: (1,1) (1,2) (2,1)`

`Class 2: (0,0) (1,0) (0,1)`

*Plot them and find the optimal separating line. What are the support vectors, and what is the*
*margin?*

```{r}
library(knitr)
library("e1071")
ABpoints <- read.table("ABpoints.tab", header=T)

model <- e1071::svm(group~., data=ABpoints, type="C-classification",
                    kernel="radial", scale = F)
```

The reason we use `scale = F` is because we want to find the scalar vector values without them having been scaled:
```{r}
support.vectors <- model$SV
knitr::kable(model$SV, format="markdown")
```

```{r}
plot(model, ABpoints)
```

We can also see the *decision boundary* ($\rho$) also known as the margin:
```{r}
model$rho
```

```{r, echo=FALSE}
# clear the environment
rm(list=ls())
```

# Problem 2
*Using the SVM R package “e1071”, Apply SVM classification to the AML vs MDS gene*
*expression datasets.*

## Dataset 207
*File “set207-MDS84.tab” contains the expression values of 25*
*key genes of 84 MDS patients*

*File “set207-AML93.tab” contains the expression values of 25*
*key genes of 93 AML patients.*

## Dataset 228
*Files for this datasets contain similar data as indicated by their file names.*

*In your analysis, use one dataset as the training set, and then apply the trained model to make*
*prediction on the other dataset.*
*Write a brief report on your analysis of the datasets: eg., Explain your choices of parameters and*
*what is the accuracy of your prediction, etc.*