---
title: "Support Vector Machine Assignment"
author: "Michael J. Jones"
date: "23/04/2015"
output: html_document
---

# Problem 1
*Suppose that the following are a set of points in two classes:*

`Class 1: (1,1) (1,2) (2,1)`

`Class 2: (0,0) (1,0) (0,1)`

*Plot them and find the optimal separating line. What are the support vectors, and what is the*
*margin?*

```{r}
library(knitr)
library("e1071")
ABpoints <- read.table("ABpoints.tab", header=T)

model <- e1071::svm(group~., data=ABpoints, type="C-classification",
                    kernel="radial", scale = F)
```

The reason we use `scale = F` is because we want to find the suppor vector values without them having been scaled:
```{r}
support.vectors <- model$SV
knitr::kable(model$SV, format="markdown")
```

```{r}
plot(model, ABpoints)
```

We can also see the *decision boundary* ($\rho$) also known as the margin:
```{r}
model$rho
```

```{r, echo=FALSE}
# clear the environment
rm(list=ls())
```

# Problem 2
*Using the SVM R package “e1071”, Apply SVM classification to the AML vs MDS gene*
*expression datasets.*

**Dataset 207**

*File “set207-MDS84.tab” contains the expression values of 25*
*key genes of 84 MDS patients*

*File “set207-AML93.tab” contains the expression values of 25*
*key genes of 93 AML patients.*

**Dataset 228**

*Files for this datasets contain similar data as indicated by their file names.*

*In your analysis, use one dataset as the training set, and then apply the trained model to make*
*prediction on the other dataset.*
*Write a brief report on your analysis of the datasets: eg., Explain your choices of parameters and*
*what is the accuracy of your prediction, etc.*

Given our datasets, let's see how many patients are in each set. We take away the first two 
columns which are the `uniqueID` and `tables_count` columns.

```{r}
num.patients <- sapply(X=dir(pattern = "set*"), FUN=function(X) {
  ncol(read.table(X)[,c(-1,-2)])
})
knitr::kable(as.data.frame(num.patients), format="markdown")
```

We will use **Dataset 207** as our training data set. In order to use the `e1071::svm` command with our data, we first need to transform our data into a data frame. The data frame will contain a row 
for each patient with columns for each of the probe IDs and a further column indicating whether the 
patient is labeled as having *AML* and *MDA*. The same will be done for loading the test data.

## Loading Dataset 207 into a dataframe
First, let's load the `set207-AML93` data set into a data.frame
```{r}
set207.AML93 <- read.table(file="set207-AML93.tab", header=T)
knitr::kable(head(set207.AML93)[,1:3], format="markdown")
```

Next, we remove `uniqueID` and `tables_count` columns, but first saving the uniqueIDs (the probeset 
IDs) in a separate vector.

```{r}
probe.ids <- as.vector(set207.AML93$uniqueID)
set207.AML93 <- set207.AML93[,-c(1,2)]
```

Having done this, we can transpose the table using the `t()` command
```{r}
set207.AML93 <- t(set207.AML93)
```

We then add an extra column to show that each of the patients is labeled as being part of the group "AML" and finally 
adding the probeset IDs and "group" as column names to the resulting dataframe.
```{r}
set207.AML93 <- data.frame("AML", set207.AML93)
colnames(set207.AML93) <- c("group", probe.ids)

knitr::kable(head(set207.AML93)[,1:4], format="markdown")
```

Further to this, we do exactly the same for `set207.MDS84` dataset. (Note we *can* use the `probe.ids` vector from before to generate 
to name the columns.)

```{r}
set207.MDS84 <- read.table(file="set207-MDS84.tab", header=T)
set207.MDS84 <- set207.MDS84[,-c(1,2)]

set207.MDS84 <- t(set207.MDS84)

set207.MDS84 <- data.frame("MDS", set207.MDS84)
colnames(set207.MDS84) <- c("group", probe.ids)

knitr::kable(head(set207.MDS84)[,1:4], format="markdown")
```

All that remains to load *Dataset 207* is to bind the `set207.AML93` dataset with `set207.MDS84`.
```{r}
set207 <- rbind(set207.AML93, set207.MDS84)
```

## Loading Dataset 228 into a dataframe
`set207` will be used to train our Support Vector Machine (SVM) such that we can give it unlabeled data points that the machine 
will classify as either being part of the `AML` or `MDS` group. Before we can do this, we need to use a testing set of data which is 
also labeled data that we can run through the algorithm and check whethe they have been classified correctly.

`set228` will be used as the testing set of data. Let's first load *Dataset 228* into a dataframe like we created above for Dataset 207.

```{r}
set228.AML109 <- read.table(file="set228-AML109.tab", header=T)
set228.AML109 <- set228.AML109[,-c(1,2)]

set228.AML109 <- t(set228.AML109)

set228.AML109 <- data.frame("AML", set228.AML109)
colnames(set228.AML109) <- c("group", probe.ids)

# TODO same for set228.MDS80...
set228.MDS80 <- read.table(file="set228-MDS80.tab", header=T)
set228.MDS80 <- set228.MDS80[,-c(1,2)]

set228.MDS80 <- t(set228.MDS80)

set228.MDS80 <- data.frame("MDS", set228.MDS80)
colnames(set228.MDS80) <- c("group", probe.ids)

set228 <- rbind(set228.AML109, set228.MDS80)
```